---
title: "Project Proposal"
format: html
editor: visual
names: Jefferson Ratliff, Evi Jonas, Sydney Stitt
---

# Proposal – Dallas Animals

```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(ggplot2)
library(dplyr)
library(readr)
```

```{r}
readr::read_csv("week18_dallas_animals.csv")%>%
  head(n = 5)
dog <- read_csv("week18_dallas_animals.csv")
```

```{r}
#Code for dataframe dimensions
print(dim(dog))
```

Data Overview

This data set uses information from a variety of different animal shelters across the City of Dallas, upon their intake to the shelter. Each observation/row is an individual animal entered by staff into the shelter system. There are 34819 observations. The data is across all animal types including dogs, cats and other wildlife.

We chose this data set because we are dog lovers and one of our group members is from Dallas and volunteered at a local shelter. It also seemed to have well collected data with a lot of diversity within key variables like dog breed and intake form.

Variables:

1\) Animal_id (ID number for each animals)

2\) Animal_type (dog, cat, wildlife)

3\) Animal_Breed

4\) Month (intake month)

5\) Month (intake Year)

6\) Census_tract (Geographical Location)

7\) Council_district (Geographical Boundries)

8\) Intake_type

9\) Outcome_type

10\) Outcome_date(When animals leave the shelter)

11\) Chip_status (is the animal chipped?)

12\) animal_origin (How they got the animal)

Discussion Questions

1.  Which city council districts have the highest rate of adoption? (council_district and outcome_type).
    1.  Filter outcome_type for observations only including adoptions using dplyr.
    2.  Quantify how many observations of adoptions there are in each district.
    3.  Graph using bar graph with ggplot2.
2.  What types of animal intake tend to stay longer in shelters?
    1.  Mutate new variable for duration of stay with dplyr
    2.  Graph using bar graph with error bars for average stay (y axis) and type of intake (x axis) with ggplot2
