---
title: "Dallas Dog - write-up"
format: html
editor: visual
names: Jefferson Ratliff, Evi Jonas, Sydney Stitt
---

```{r}
# Do not modify this chunk.
install.packages("tmap")

knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(ggplot2)
library(dplyr)
library(readr)
library(lubridate)
library(tmap)
library(sf)
```

Introduction:

Our data set we chose collected data across a period of one year between 2016-2017. The data is collected to keep citizens informed about the operation and activities of the shelters across the city. This data set uses information from a variety of different animal shelters across the City of Dallas, upon their intake to the shelter. Each observation/row is an individual animal entered by staff into the shelter system. There are 34819 observations. The data is across all animal types including dogs, cats and other wildlife.

We chose this data set because we are dog lovers and one of our group members is from Dallas and volunteered at a local shelter. It also seemed to have well collected data with a lot of diversity within key variables like dog breed and intake form.

**Question 1:** Which city council districts have the highest rate of adoption?

Introduction:

The question at hand involves investigating the adoption rates across various city council districts, specifically focusing on the variables council_district and outcome_type. To address this query effectively, it is crucial to filter the data set using the dplyr package to isolate observations related to adoptions. By narrowing down the data set to only include these instances, we can then proceed to quantify the number of adoptions in each city council district. To visually represent this information, the ggplot2 package will be employed to create a bar graph, offering a clear and concise presentation of the adoption rates across different districts.

To accomplish this analysis, the data set needs to include relevant information such as council_district and outcome_type. The data set should be pre-processed using dplyr functions to filter and extract the necessary observations related to adoptions. The quantification process involves counting the number of adoption instances within each city council district. Finally, the visualization aspect will be addressed using ggplot2 to generate a bar graph that illustrates the variations in adoption rates across different districts. My interest in this question is driven by the need to understand the geographical distribution of adoption rates within a city, which can have implications for resource allocation, community outreach, and policy planning to enhance animal welfare and adoption initiatives.

Approach:

To comprehensively address the question of dog adoption rates across city council districts, a combination of visualizations has been employed. The initial bar plot succinctly illustrates the raw counts of adoptions in each district, providing a quick overview of the districts with higher adoption rates. Building upon this, a box plot offers a more detailed perspective, showcasing the distribution of adoption rates within each district. This plot is instrumental in identifying any potential variations, outliers, or skewness in the adoption data. Finally, a faceted bar plot takes the analysis a step further by presenting individual bar plots for each council district. This approach allows for a nuanced examination of adoption patterns in each district while maintaining a clear comparative framework. Together, these visualizations offer a comprehensive understanding of the adoption landscape, enabling insights into both the overall trends and the variability within specific city council districts.

Analysis:

```{r}
#finding number of total adoptions per district in the data frame
dog_adoptions <- dog %>%
  filter(outcome_type == "ADOPTION") %>% 
  group_by(council_district) %>% 
  summarize(
    n_adoptions = n()
  )
```

```{r}
# Now let's create the plot
ggplot(data = dog_adoptions, aes(x = council_district, y = n_adoptions)) +
  geom_bar(stat = "identity") +
  labs(x = "Council District", y = "Number of Adoptions") +
  ggtitle("Dog Adoptions by Council District")
```

Discussion:

There appears to be a wide array of differences in of adoptions by council district with districts 1, 4, 6, and 8 making up the largest four. Districts 10-14 make up the least amount of adoptions. We were able to find such a wide difference by council district reveling to us that geography is a major factor in pet adoption in the City of Dallas. Our visualization provides a clear understanding and answer to our research question.

**Question 2:** What types of animal intake tend to stay longer in shelters?

Introduction:

Approach:

Analysis:

```{r}
#Making a duration variable using dplyr
dog_duration <- dog %>%
   mutate(intake_date = mdy(intake_date),
         outcome_date = mdy(outcome_date)) %>%
  mutate(duration = difftime(outcome_date, intake_date, units = "days")) %>%
  mutate(duration = as.vector(duration))

#Getting summary statistics for average duration
dog_duration_summary <- dog_duration %>%
  group_by(intake_type) %>%
  summarize(average_duration = mean(duration, na.rm = TRUE),
            sd_duration = sd(duration, na.rm = TRUE),
            count_types = n()) 

#Graphing violin plot (all data)
ggplot(dog_duration, aes(x = intake_type, y = duration)) +
  geom_violin() +
  theme_minimal(base_size = 8) +
  labs(x = "Intake Type", 
       y = "Duration of Stay", 
       title = "Shelter Intake Types and Durations in Dallas Animal Shelters in 2016 and 2017")

#Graphing bar graph, sd is larger than mean which means there is abnormal distributions
ggplot(dog_duration_summary) +
  geom_bar(aes(x = intake_type, y = average_duration), 
           stat = "identity") +
  geom_errorbar(aes(x = intake_type, 
                    ymin = average_duration - sd_duration, 
                    ymax = average_duration + sd_duration)) +
    labs(x = "Intake Type", 
       y = "Duration of Stay with Error Bars", 
       title = "Shelter Intake Types and Durations in Dallas Animal Shelters in 2016 and 2017")

#Graphing histograms, y scale is messed up because stray has many more observations than other groups
ggplot(dog_duration, aes(x = duration)) +
  geom_histogram() +
  facet_wrap(~intake_type) +
  labs(x = "Intake Type", 
       y = "Duration of Stay", 
       title = "Shelter Intake Duration Distrib in Dallas Animal Shelters in 2016 and 2017")
```

Discussion:
